#Lab_03 {lab_03}

rob wells, 8/14/2021

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
options(repos = "https://cloud.r-project.org") 
library(tidyverse)

# Load janitor here
library(janitor)

```

### Load data

In this data set, each row represents a city or town in Maryland with the 2010 and 2020 figures for:

-   Total population
-   White only population
-   Black or African American only population
-   Hispanic only population
-   Native American only population
-   Asian only population
-   Pacific Islander (including native Hawiian) only population

Background on this data from [Social Explorer](https://www.socialexplorer.com/data/ACS2016_5yr/metadata/?ds=ACS16_5yr&table=B02001)

```{r}
# Write code to load the Maryland race by city, place for 2010 - 2020 and call it md_race.
# The data is called lab_03.csv, and is in the same folder as lab_03.qmd (this data notebook)

#View(lab_03)

```

## Answer questions

Q1: Which five cities have the largest overall Hispanic population in 2020? Name the five cities that had the largest Hispanic population in 2020.

A1:

```{r}

# Write your code in this codeblock.
lab_03 %>% 
  select(place,x2020_hispanic) %>% 
  slice_max(x2020_hispanic, n = 5)
#The cities that had the largest Hispanic Population in 2020 were: Baltimore City, Wheaton CDP, Germantown CDP, Silver Spring CDP and Aspen Hill CDP. 
```

Q2: Find the cities with the largest Black population and arrange the results in descending order.

A2:

```{r}
lab_03 %>% 
  select(place,x2020_black) %>% 
  arrange(desc(x2020_black))

#The top 5 cities with the largest black Population is Baltimore City, Waldorf CDP, Bowie City, Clinton CDP and Colombia CDP. 

```

Q3: Create a table called top_hispanic that shows the top 15 communities with the largest Hispanic population in 2020, and also display the 2010 Hispanic population and the total community population for 2020. Look at your results - what stands out to you? Craft a one-paragraph story pitch for your editor.

A3:

```{r}
lab_03 %>% 
  select(place,x2010_hispanic,x2020_hispanic,x2020_total) %>% 
  slice_max(x2020_hispanic, n = 15)

# What stood out to me the most was that Langley Park CDP's Hispanic population makes up about 83% of the cities overall population. I would want to explore why the Hispanic population is the majority there and if it in increasing Hispanic population in cities near by. Silver Spring is close to Langley Park but the percent of Hispanic population is much lower. I would be interested to look closely at the population of Langley park and white the Hispanic population is most prominent there. I also could expand my search into the county Hispanic population of Langley Park. 

```
