#lab_11
Rob Wells, April 11, 2024

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Overview

There are three main goals of this lab:

* learn to make use of R packages that can access data via third-party APIs.
* learn to develop a quick and accurate understanding of what data can be obtained via third-party APIs. 
* learn to develop questions that can be answered using that data, and avoid developing questions that cannot. 

## APIs to choose from
 

* tidycensus - for asking census-related questions.

## Load libraries

**Task**: load the tidyverse and janitor.  Uncomment the one other package you plan to work with.  Do not choose more than one API package.

```{r}
library(tidyverse)
library(janitor)
library(tidycensus)

```

## Get setup to access third-party APIs

### tidycensus

Use the U.S. census access token you created for this week's lab.  You can store it by editing the block below to replace 'YOUR KEY HERE' with your actual key, a long string of letters and numbers. If you need to sign up for another one, you can do so here:  [apply for an API key with the Census Bureau](https://api.census.gov/data/key_signup.html). 

```{r}
census_api_key( "8be73b673e80b7654a9b35fbb31bbd53774f1706", install=TRUE, overwrite = TRUE)
```



### tidycensus

This package is designed to help you pull demographic information about different U.S. geographies.

* Basic usage https://walker-data.com/tidycensus/articles/basic-usage.html
* Functions: https://walker-data.com/tidycensus/reference/index.html
* More details https://walker-data.com/tidycensus/

Sample question: which state saw the largest growth in Hispanic population between 2010 and 2020?

```{r}

acs_vars <- load_variables( dataset = "acs5", year = 2020) 
  
 fem <- get_acs(geography = "state",
              variables = c(female_population = "B01001_026"),
              year = 2020) 
  pop <- get_acs(geography = "state",
              variables = c(total_population = "B01001_001"),
              year = 2020) 
 
 join_pct <- inner_join(fem, pop, by = c( "NAME")) %>%
   mutate(pct_whole = (estimate / total_population) *100 )%>%
  arrange(desc(pct_whole))
```

### Develop and answer questions

**Task** Develop two different *interesting* questions that you can answer with the data. Run them by your professor before getting started on answering, to ensure they're sufficiently interesting and possible with the data. 
```{r}
# 1. What state has the largest and smallest population of women in the U.S.?
# 2. How many white women in there 20s live in New York 
```

**Task** Write code to answer the questions.  Write up a sentence with each of your findings. Produce at least one visualization to support one of your findings. 
```{r}
 # I want to see where I should live if I want to have female roomantes after college!!!

 fem <- get_acs(geography = "state",
              variables = c(female_population = "B01001_026"),
              year = 2020) %>% 
              mutate(fem_estimate = estimate)

  pop <- get_acs(geography = "state",
              variables = c(total_population = "B01001_001"),
              year = 2020) %>% 
              mutate(total_estimate = estimate)
 
 join_pct <- fem %>% 
   inner_join(pop, by = c("NAME"))%>%
   mutate(pct_whole = ( fem_estimate / total_estimate) *100 )%>%
  arrange(desc(pct_whole))%>%
select(NAME, fem_estimate, total_estimate, pct_whole )%>%
   arrange(desc(pct_whole))
 
#D.C has the largest population of women out of the total percentage with a pct_whole of 52.5%

```

```{r}
 NY <- get_acs(geography = "county",
              variables = c(population_of_women_20_24 = "B01001A_023", population_of_women_25_29 = "B01001A_024"),
              state = "NY",
              year = 2020) 

NY <- NY %>%
sum(NY$estimate)

#the total amount of women in their 20s that live in NY is 792681. 
```

