#lab_08
Sean Mussenden, Aug 16, 2022

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load janitor and the tidyverse
```{r include=FALSE}
#background on tidyverse: https://www.tidyverse.org/packages/
# install.packages("formattable")
#install.packages("janitor")
#install.packages("tidyverse")
#install.packages("jsonlite")
library(tidyverse)
library(janitor)
library(lubridate)
library(jsonlite)
```

## Load data
```{r include=FALSE}
library(readr)
total_pop <- read_csv("~/Documents/GitHub/data_journalism_class/04_labs/lab_08/data/maryland_population.csv")
hispanic_pop <- read_csv("~/Documents/GitHub/data_journalism_class/04_labs/lab_08/data/maryland_hispanic.csv")
white_pop <- read_csv("~/Documents/GitHub/data_journalism_class/04_labs/lab_08/data/maryland_white.csv")
black_pop <- read_csv("~/Documents/GitHub/data_journalism_class/04_labs/lab_08/data/maryland_black.csv")
```
**Task**: Load four tables of U.S. Census data.  Each table has one row per Maryland county with 2020 population information.  Each table has these same two columns:

* geoid - a unique identifier for each county
* name - the county name

The third column in each table represents the total population in one of four demographic categories.

* maryland_population.csv - total_pop - total population 
* maryland_hispanic.csv - hispanic_pop - total Hispanic population
* maryland_white.csv - white_pop - total White alone population
* maryland_black.csv - black_pop - total Black alone population

Some of the information in the tables may need to be cleaned or validated in order to produce the correct answer.  

## Questions

**Q1**  What Maryland county has the largest Hispanic population (by percentage of total population)? 
**A1** 
```{r include=FALSE}
hispanic_pop <- hispanic_pop  %>%
  mutate(name = case_when(name == "Montgomery County,Maryland" ~ "Montgomery County, Maryland", TRUE ~name)) 

md_master <- total_pop %>%
inner_join(black_pop)%>%
inner_join(white_pop) %>% 
inner_join(hispanic_pop)  %>%
arrange(desc(hispanic_pop))
  
#the Maryland county with the largest Hispanic community is Montgomery County. 
```

**Q2**  What is the total population of people who identify as white alone in Maryland? 
**A2**  
```{r include=FALSE}
#Somerset's white population of 13510000 needs to be divided by 100

white_pop <- white_pop %>%
  mutate(white_pop = case_when(
    white_pop == "13510000" ~ 13510000/100,
    TRUE ~ white_pop
  ))

md_master <- md_master %>%
  inner_join(black_pop) %>% 
  inner_join(white_pop) 
  #the total population of people who identify as white alone in Maryland is. 
```
**Q3** What is the average Black population in all Maryland counties? 
**A3** 
```{r include=FALSE}
#Link https://www.census.gov/quickfacts/fact/table/baltimorecitymaryland/RHI225222#RHI225222 
# Baltimore City's April 1, 2020 population was 585,708 and 61.2% of the population was Black,358,453 people.
#Baltimore County's April 1, 2020 population was 854,535 and 31.9% of the population was Black, so this is 272,596 people.

md_black <- black_pop %>% 
  mutate(black_pop = (case_when(
    name == "Baltimore County, Maryland" ~ 272596, 
    name == "Baltimore city, Maryland" ~ 358453,
    TRUE ~ black_pop
)))
```
